---
title: "gtsummary in quarto"
format: 
 html:
  theme: yeti
  execute:
  echo: false
---

```{r}
#| warning: false
library(tidyverse)
library(gtsummary)
nlsy_cols <- c("glasses", "eyesight", "sleep_wkdy", "sleep_wknd", 
               "id", "nsibs", "samp", "race_eth", "sex", "region", 
               "income", "res_1980", "res_2002", "age_bir")
nlsy <- read_csv(here::here("data", "raw", "nlsy.csv"), 
                 na = c("-1", "-2", "-3", "-4", "-5", "-998"),
                 skip = 1, col_names = nlsy_cols) |> 
  mutate(region_cat = factor(region, labels = c("Northeast", "North Central", "South", "West")),
         sex_cat = factor(sex, labels = c("Male", "Female")),
         race_eth_cat = factor(race_eth, labels = c("Hispanic", "Black", "Non-Black, Non-Hispanic")),
         eyesight_cat = factor(eyesight, labels = c("Excellent", "Very good", "Good", "Fair", "Poor")),
         glasses_cat = factor(glasses, labels = c("No", "Yes"))) 
```

# Customization of `tbl_summary()`

```{r}
mean_weekend<-  mean(nlsy$sleep_wknd, na.rm=TRUE)
mean_weekend
```
Average sleep on weekends was `r mean_weekend`

```{r}
#| label: tbl-overall
#| tbl-cap: "Overll Statistics"
initial_table <-tbl_summary(
  nlsy,
  include = c(sex_cat, race_eth_cat, region_cat,
              eyesight_cat, glasses, age_bir))

initial_table
```

```{r}
#| label: tbl-sex
#| tbl-cap: "Descriptive statistics"
table <- tbl_summary(
  nlsy,
  by = sex_cat,
  include = c(sex_cat, race_eth_cat, region_cat,
              eyesight_cat, glasses, age_bir, income),
  label = list(
    race_eth_cat ~ "Race/ethnicity",
    region_cat ~ "Region",
    eyesight_cat ~ "Eyesight",
    glasses ~ "Wears glasses",
    age_bir ~ "Age at first birth"
  ),
  missing_text = "Missing") |>
 add_overall()
table
```

```{r}
glassesf<-inline_text(table, variable = "glasses", column = "stat_1",
            pattern = "{n}/{N} ({p}%)")
glassesm<-inline_text(table, variable = "glasses", column = "stat_1",
            pattern = "{n}/{N} ({p}%)")
```
`r glassesf` number of Females wore glasses.
`r glassesm` number of males wore glasses.

```{r}
income_f <- inline_text(table, variable = "income", 
                         column = "stat_2")
income_m <- inline_text(table, variable = "income", 
                         column = "stat_1")
```
A greater proportion of female (`r income_f`) than male (`r income_m`) participants wore glasses.

```{r}
#| label: tbl-descr
#| tbl-cap: "Descriptive statistics by sex "
table5<-tbl_summary(
  nlsy,
  by = sex_cat,
  include = c(sex_cat, race_eth_cat, region_cat,
              eyesight_cat, glasses, age_bir))
table5
```

This is (@tbl-sex) and (@tbl-descr).

```{r}
table90 <- tbl_summary(
  nlsy,
  by = sex_cat,
  include = c(sex_cat, race_eth_cat, region_cat,
              eyesight_cat, glasses, age_bir, sleep_wknd, sleep_wkdy))
```

```{r}
sleepf <- inline_text(table90, variable = "sleep_wknd", 
                         column = "stat_2")
sleepm <- inline_text(table90, variable = "sleep_wknd", 
                         column = "stat_1")

```

Females had `r sleepf` hours of sleep and males had `r sleepm` hours of sleep.



```{r}
#| label: tbl-race
#| tbl-cap: "Racial statistics"
table1 <- tbl_summary(
  nlsy,
  by = race_eth_cat,
  include = c(sex_cat, race_eth_cat, region_cat,
              eyesight_cat, glasses, age_bir)) |> 
  add_overall(last = TRUE)
table1
```

```{r}
glassesh <- inline_text(table1, variable = "glasses", 
                         column = "stat_1")
glassesb <- inline_text(table1, variable = "glasses", 
                         column = "stat_2")
glassesnb <- inline_text(table1, variable = "glasses", 
                         column = "stat_3")
```

We observed that `r glassesh` of hispanic individuals wore glasses. 

We observed that `r glassesb` of black individuals wore glasses. 

We observed that `r glassesnb` of non-black/non-hispanic individuals wore glasses. 
